---
title: "ANALYSIS OF VARIOUS BDD TYPES"
author: "Robert Meolic"
date: '`r format(Sys.Date(), "%d\\. %m\\. %Y")`'
output:
  pdf_document: default
  html_document: default
---

Calculated with Biddy (biddy.meolic.com).

Raw data are available in the SVN repository.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
options(width=80)
options(scipen=999)
```

```{r, echo=FALSE}
OBDD <- "Ordered Binary Decision Diagram"
OBDDCE <- "Ordered Binary Decision Diagram with complemented edges"
ZBDD <- "Zero-suppressed Binary Decision Diagram"
ZBDDCE <- "Zero-suppressed Binary Decision Diagram with complemented edges"
TZBDD <- "Tagged Zero-suppressed Binary Decision Diagram"
TZBDDCE <- "Tagged Zero-suppressed Binary Decision Diagram with complemented edges"
```

```{r, echo=FALSE}
OBDDCOLOR <- "blue4"
OBDDCECOLOR <- "blue1"
ZBDDCOLOR <- "green4"
ZBDDCECOLOR <- "green1"
TZBDDCOLOR <- "brown4"
TZBDDCECOLOR <- "brown1"
```

```{r, echo=FALSE}
OBDDSTYLE <- 1
OBDDCESTYLE <- 1
ZBDDSTYLE <- 5
ZBDDCESTYLE <- 5
TZBDDSTYLE <- 4
TZBDDCESTYLE <- 4
```

```{r, echo=FALSE}
for (TYPE in c("OBDD","OBDDCE","ZBDD","ZBDDCE")) {
  for (SIZE in c(2,3,4,5)) {
    NAME <- paste("SUM.",TYPE,".0",SIZE,sep="")
    FULLNAME <- paste(NAME,"FULL",sep=".")
    FILENAME <- paste("./","SUM-",TYPE,"-0",SIZE,".TXT",sep="")
    TABLE <- read.table(FILENAME, header=TRUE, quote="\"")
    data.table::setnames(TABLE,"NODES", "N")
    assign(FULLNAME,TABLE)
    IND <- (TABLE[[5]] != 0)
    TABLE <- TABLE[IND, ]
    TABLE$CMIN <- cumsum(as.numeric(TABLE$MIN))
    TABLE$CMAX <- cumsum(as.numeric(TABLE$MAX))
    TABLE$CBASE <- cumsum(as.numeric(TABLE$BASE))
    TABLE$CALL <- cumsum(as.numeric(TABLE$ALL))
    TABLE$RMIN <- formattable::percent(TABLE$CMIN / max(TABLE$CMIN))
    TABLE$RMAX <- formattable::percent(TABLE$CMAX / max(TABLE$CMAX))
    TABLE$RBASE <- formattable::percent(TABLE$CBASE / max(TABLE$CBASE))
    TABLE$RALL <- formattable::percent(TABLE$CALL / max(TABLE$CALL))
    assign(NAME,TABLE)
    assign(paste("SIZE",TYPE,SIZE,sep=""),length(TABLE$N))
  }
}
```

# USED ACRONYMS

```{r, echo=FALSE, results='asis'}
for (TYPE in c("OBDD","OBDDCE","ZBDD","ZBDDCE")) {
  cat(TYPE)
  cat(" - ")
  cat(get(TYPE))
  cat("\n\n")
}
```

MIN - Number of Boolean functions with n nodes using the optimal ordering

MAX - Number of Boolean functions with n nodes using the worst ordering

BASE - Number of Boolean functions with n nodes using the base ordering

ALL - Number of BDDs with n nodes for all Boolean functions for all orderings

```{r, echo=FALSE, fig.align='center', results='asis'}
# thanks to Andrew Tyre for help with R
# https://stackoverflow.com/questions/49560705/use-r-to-calculate-median-without-replicating-elements
MIN <- "using the optimal ordering (MIN)"
MAX <- "using the worst ordering (MAX)"
BASE <- "using the base ordering (BASE)"
for (TYPE in c("OBDD","OBDDCE","ZBDD","ZBDDCE")) {
  for (SIZE in c(2,3,4,5)) {
    NAME <- paste("SUM.",TYPE,".0",SIZE,sep="")
    TABLE <- get(NAME)
    STABLE <- data.frame(STAT = c("mean","median","mode"))
    STABLE$MIN[1] <- Hmisc::wtd.mean(TABLE$N, weights = as.numeric(TABLE$MIN))
    STABLE$MIN[2] <- Hmisc::wtd.quantile(TABLE$N, probs = c(0.5), weights = as.numeric(TABLE$MIN))
    STABLE$MIN[3] <- TABLE$N[which.max(TABLE$MIN)]
    STABLE$MAX[1] <- Hmisc::wtd.mean(TABLE$N, weights = as.numeric(TABLE$MAX))
    STABLE$MAX[2] <- Hmisc::wtd.quantile(TABLE$N, probs = c(0.5), weights = as.numeric(TABLE$MAX))
    STABLE$MAX[3] <- TABLE$N[which.max(TABLE$MAX)]
    STABLE$BASE[1] <- Hmisc::wtd.mean(TABLE$N, weights = as.numeric(TABLE$BASE))
    STABLE$BASE[2] <- Hmisc::wtd.quantile(TABLE$N, probs = c(0.5), weights = as.numeric(TABLE$BASE))
    STABLE$BASE[3] <- TABLE$N[which.max(TABLE$BASE)]
    PERCENTILES <- c(0.1, 0.25, 0.5, 0.75, 0.9)
    PTABLE <- data.frame(PERCENTILE = formattable::percent(PERCENTILES,digits=0))
    PTABLE$MIN <- Hmisc::wtd.quantile(TABLE$N, probs = PERCENTILES, weights = as.numeric(TABLE$MIN))
    PTABLE$MAX <- Hmisc::wtd.quantile(TABLE$N, probs = PERCENTILES, weights = as.numeric(TABLE$MAX))
    PTABLE$BASE <- Hmisc::wtd.quantile(TABLE$N, probs = PERCENTILES, weights = as.numeric(TABLE$BASE))
    cat("\\pagebreak\n\n")
    cat("#")
    cat(TYPE)
    cat(" for Boolean functions with ",SIZE," variables\n\n")
    print(knitr::kable(TABLE[c(1,2,10,3,11,4,12,5,13)],row.names=FALSE))
    cat("\n\n\\begin{center}\\begin{tabular}{p{0.3\\linewidth}p{0.1\\linewidth}p{0.3\\linewidth}}\n")
    print(knitr::kable(PTABLE,row.names=FALSE,format="latex",booktabs=TRUE))
    cat("\n&&\n")
    print(knitr::kable(STABLE,row.names=FALSE,format.args=list(nsmall=2),format="latex",booktabs=TRUE))
    cat("\\end{tabular}\\end{center}\\bigskip")
    plot(TABLE$N,TABLE$N,ylim=c(0,1),cex.axis=0.8,cex.lab=0.8,lab=c(length(TABLE$N),8,0),mgp=c(2,1,0),xlab="Number of nodes",ylab="Fraction of realizable Boolean functions",type="n")
    lines(TABLE$N,TABLE$RMIN,type="o", col="green")
    lines(TABLE$N,TABLE$RMAX,type="o", col="red")
    lines(TABLE$N,TABLE$RBASE,type="o", col="blue")
    abline(h=0.0, lty=3, col="gray40")
    abline(h=0.25, lty=3, col="gray40")
    abline(h=0.5, lty=3, col="gray40")
    abline(h=0.75, lty=3, col="gray40")
    abline(h=1.0, lty=3, col="gray40")
    legend(1,legend=c(MIN,MAX,BASE),col=c("green","red","blue"),lty=c(1,1,1),cex=0.8)
    cat("\n\n")
  }
}
```

\pagebreak

# BDD TYPE COMPARISONS

```{r, echo=FALSE, results='asis'}
RMIN <- "using the optimal ordering (MIN)"
RMAX <- "using the worst ordering (MAX)"
RBASE <- "using the base ordering (BASE)"
for (PARAM in c("RMIN","RMAX","RBASE")) {
  for (SIZE in c(2,3,4,5)) {
    N <- max(get(paste("SIZEOBDD",SIZE,sep="")),get(paste("SIZEZBDD",SIZE,sep="")))
    plot(0,xlim=c(1,N),ylim=c(0,1),cex.axis=0.8,cex.lab=0.8,lab=c(N,8,0),xlab="Number of nodes",ylab="",mgp=c(2,1,0),type="n")
    abline(h=0.0, lty=3, col="gray40")
    abline(h=0.25, lty=3, col="gray40")
    abline(h=0.5, lty=3, col="gray40")
    abline(h=0.75, lty=3, col="gray40")
    abline(h=1.0, lty=3, col="gray40")
    TT <- c()
    CC <- c()
    SS <- c()
    for (TYPE in c("OBDD","OBDDCE","ZBDD","ZBDDCE")) {
      NAME <- paste("SUM.",TYPE,".0",SIZE,sep="")
      #cat("\n\n\\pagebreak\n\n")
      TABLE <- get(NAME)
      C <- get(paste(TYPE,"COLOR",sep=""))
      S <- get(paste(TYPE,"STYLE",sep=""))
      lines(TABLE$N,TABLE[[PARAM]],type="o",lty=S,col=C)
      TT <- append(TT,TYPE)
      CC <- append(CC,C)
      SS <- append(SS,S)
    }
    title(paste("Fraction of realizable Boolean functions with ",SIZE," variables ",get(PARAM),sep=""),cex.main=0.8,line=1)
    legend(1,legend=TT,col=CC,lty=SS,cex=0.8)
    cat("\n\n\\bigskip\n\n")
  }
}
```

```{r, echo=FALSE, results='asis'}
RMIN <- "using the optimal ordering (MIN)"
RMAX <- "using the worst ordering (MAX)"
RBASE <- "using the base ordering (BASE)"
for (PARAM in c("RMIN","RMAX","RBASE")) {
  plot(0,xlim=c(1,N),ylim=c(0,1),cex.axis=0.8,cex.main=0.8,,cex.lab=0.8,lab=c(N,8,0),xlab="Number of nodes",ylab="",mgp=c(2,1,0),type="n")
  abline(h=0.0, lty=3, col="gray40")
  abline(h=0.25, lty=3, col="gray40")
  abline(h=0.5, lty=3, col="gray40")
  abline(h=0.75, lty=3, col="gray40")
  abline(h=1.0, lty=3, col="gray40")
  for (SIZE in c(2,3,4,5)) {
    N <- max(get(paste("SIZEOBDD",SIZE,sep="")),get(paste("SIZEZBDD",SIZE,sep="")))
    TT <- c()
    CC <- c()
    SS <- c()
    for (TYPE in c("OBDD","ZBDD")) {
      NAME <- paste("SUM.",TYPE,".0",SIZE,sep="")
      #cat("\n\n\\pagebreak\n\n")
      TABLE <- get(NAME)
      C <- get(paste(TYPE,"COLOR",sep=""))
      S <- get(paste(TYPE,"STYLE",sep=""))
      lines(TABLE$N,TABLE[[PARAM]],type="o", lty=S, col=C)
      TT <- append(TT,TYPE)
      CC <- append(CC,C)
      SS <- append(SS,S)
    }
  }
  title(paste("Fraction of realizable Boolean functions ",get(PARAM),sep=""),cex.main=0.8,line=1)
  legend("bottomright",legend=TT,col=CC,lty=SS,cex=0.8)
  cat("\n\n\\bigskip\n\n")
}
```

```{r, echo=FALSE, results='asis'}
RMIN <- "using the optimal ordering (MIN)"
RMAX <- "using the worst ordering (MAX)"
RBASE <- "using the base ordering (BASE)"
for (PARAM in c("RMIN","RMAX","RBASE")) {
  plot(0,xlim=c(1,N),ylim=c(0,1),cex.axis=0.8,cex.main=0.8,,cex.lab=0.8,lab=c(N,8,0),xlab="Number of nodes",ylab="",mgp=c(2,1,0),type="n")
  abline(h=0.0, lty=3, col="gray40")
  abline(h=0.25, lty=3, col="gray40")
  abline(h=0.5, lty=3, col="gray40")
  abline(h=0.75, lty=3, col="gray40")
  abline(h=1.0, lty=3, col="gray40")
  for (SIZE in c(2,3,4,5)) {
    N <- max(get(paste("SIZEOBDD",SIZE,sep="")),get(paste("SIZEZBDD",SIZE,sep="")))
    TT <- c()
    CC <- c()
    SS <- c()
    for (TYPE in c("OBDDCE","ZBDDCE")) {
      NAME <- paste("SUM.",TYPE,".0",SIZE,sep="")
      #cat("\n\n\\pagebreak\n\n")
      TABLE <- get(NAME)
      C <- get(paste(TYPE,"COLOR",sep=""))
      S <- get(paste(TYPE,"STYLE",sep=""))
      lines(TABLE$N,TABLE[[PARAM]],type="o", lty=S, col=C)
      TT <- append(TT,TYPE)
      CC <- append(CC,C)
      SS <- append(SS,S)
    }
  }
  title(paste("Fraction of realizable Boolean functions ",get(PARAM),sep=""),cex.main=0.8,line=1)
  legend("bottomright",legend=TT,col=CC,lty=SS,cex=0.8)
  cat("\n\n\\bigskip\n\n")
}
```
